Assessing exposure to Atlantic Basin tropical storms in United States counties

### 2017-02-05 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Re-run everything with latest versions of `countyweather` and `hurricaneexposure`

```{r compare_rain_ex_counties, echo = FALSE, fig.width = 10, fig.height = 10, efig.align = "center", fig.cap = "Comparison of rain estimates for example counties."}
plot_county_rain_compare <- function(ex_fips, ex_dir, ex_title, 
                                     get_data = FALSE){
        if(get_data){
                daily_timeseries(ex_fips, coverage = 0,
                                  date_min = "1988-01-01",
                                  date_max = "2011-12-31",
                                  var = "PRCP",
                                  out_directory = ex_dir,
                                  keep_map = FALSE)
        }

@@

        ex_weather <- readRDS(paste0(ex_dir, ex_fips, ".rds"))
        ex_weather <- as.data.frame(ex_weather$daily_data)

@@

```{r echo = FALSE, eval = FALSE}
allison_gauge <- read.csv("MWR_gauge_rain_data/allison_mwr.csv") 
allison_fips <- sprintf("%05s", as.character(unique(allison_gauge$fips)))
# allison_fips <- sample(allison_fips, 10)
date_range <- closest_dist %>% 
        filter(storm_id == "Allison-2001") %>%
        summarize(earliest_date = min(ymd(closest_date)) - ddays(2),
                  latest_date = max(ymd(closest_date)) + ddays(2))
daily_timeseries(allison_fips, coverage = 0,
                  date_min = format(date_range[1, 1], "%Y-%m-%d"),
                  date_max = format(date_range[1, 2], "%Y-%m-%d"),
                  var = "PRCP", out_directory = "county_ave/allison/",
                 keep_map = FALSE)

load("MWR_gauge_rain_data/frances_gauge.RData")
frances_fips <- sprintf("%05s", as.character(unique(frances_gauge$fips)))
# frances_fips <- sample(frances_fips, 10)
date_range <- closest_dist %>% 
        filter(storm_id == "Frances-2004") %>%
        summarize(earliest_date = min(ymd(closest_date)) - ddays(2),
                  latest_date = max(ymd(closest_date)) + ddays(2))
daily_timeseries(frances_fips, coverage = 0,
                  date_min = format(date_range[1, 1], "%Y-%m-%d"),
                  date_max = format(date_range[1, 2], "%Y-%m-%d"),
                  var = "PRCP", out_directory = "county_ave/frances/",
                 keep_map = FALSE)

load("MWR_gauge_rain_data/ike_gauge.RData")
ike_fips <- sprintf("%05s", as.character(unique(ike_gauge$fips)))
# ike_fips <- sample(ike_fips, 10)
date_range <- closest_dist %>% 
        filter(storm_id == "Ike-2008") %>%
        summarize(earliest_date = min(ymd(closest_date)) - ddays(2),
                  latest_date = max(ymd(closest_date)) + ddays(2))
daily_timeseries(ike_fips, coverage = 0,
                  date_min = format(date_range[1, 1], "%Y-%m-%d"),
                  date_max = format(date_range[1, 2], "%Y-%m-%d"),
                  var = "PRCP", out_directory = "county_ave/ike/",
                 keep_map = FALSE)

load("MWR_gauge_rain_data/alex_gauge.RData")
alex_fips <- sprintf("%05s", as.character(unique(alex_gauge$fips)))
# alex_fips <- sample(alex_fips, 10)
date_range <- closest_dist %>% 
        filter(storm_id == "Alex-2010") %>%
        summarize(earliest_date = min(ymd(closest_date)) - ddays(2),
                  latest_date = max(ymd(closest_date)) + ddays(2))
daily_timeseries(alex_fips, coverage = 0,
                  date_min = format(date_range[1, 1], "%Y-%m-%d"),
                  date_max = format(date_range[1, 2], "%Y-%m-%d"),
                  var = "PRCP", out_directory = "county_ave/alex/",
                 keep_map = FALSE)
```

```{r county_ave_comparison, echo = FALSE, fig.width = 10, fig.height = 10, fig.align = "center", fig.cap = "Comparison of county-level average cumulative rainfall for the period from two days before to one day after the storm's closest approach to a county, using NLDAS-2 versus ground-based monitors, for counties with both estimates available for four example storms. The size of each point indicates the average number of monitors reporting from that county during the window (larger size indicates more monitors). The text over each point gives the FIPS county identifier."}
plot_county_ave_compare <- function(county_dir, storm_id){
        storm_title <- paste0(paste(unlist(strsplit(
                as.character(storm_id), "-")),
                             collapse = " ("), ")")
        storm_year <- gsub(".+-", "", storm_id)
        counties <- gsub(".rds", "",
                         list.files(paste0("county_ave/", county_dir)))
        if(county_dir == "frances"){
                counties <- counties[counties != "51019"] # Missing all data
        }

@@

        for(x in counties){
                county_weather <- readRDS(paste0("county_ave/", 
                                                 county_dir, "/", x,
                                                 ".rds"))$daily_data %>%
                        dplyr::mutate(date = ymd(date),
                                      prcp = prcp / 10) %>%
                        dplyr::mutate(fips = x)
                storm_ave[[x]] <- county_weather
                }
        storm_ave <- do.call("rbind", storm_ave)

@@

        compare_wind <- county_wind(counties = unique(data$fips),
                             start_year = storm_year, 
                             end_year = storm_year,
                             wind_limit = 0,
                             wind_var = "max_sust") %>%
                dplyr::filter(storm_id == data$storm_id[1]) %>%
                dplyr::right_join(data, by = "fips")

        county_range <- compare_wind %>% 
                group_by(fips) %>%
                summarize(lowest_wind = min(wind_sust),
                  highest_wind = max(wind_sust),
                  model_wind = mean(max_sust))

        plot_mwr <- ggplot(compare_wind, 
                   aes(x = max_sust, y = wind_sust)) + 
        geom_hline(yintercept = 15, color = "lightgray") + 
        geom_vline(xintercept = 15, color = "lightgray") + 
                geom_abline(aes(intercept = 0, slope = 1), color = "gray") + 
        geom_segment(data = county_range, 
                     aes(x = model_wind, xend = model_wind,
                         y = lowest_wind, yend = highest_wind,
                         group = fips)) + 
        geom_point(aes(group = NA), alpha = 0.5) + 
        ggthemes::theme_few() + 
        xlim(range(compare_wind$max_sust, compare_wind$wind_sust)) +
        ylim(range(compare_wind$max_sust, compare_wind$wind_sust)) + 
        ggtitle(storm_title) + 
        xlab("Modeled sustained wind (m / s)") + 
        ylab("Monthly Weather Report\n sustained wind reports (m / s)") + 
        scale_color_manual(values = c("TRUE" = "blue", "FALSE" = "black"), 
                           guide = "none") 
        return(plot_mwr)
}

@@

```{r top_wind, echo = FALSE, results = "asis"}
a <- wind_exposure %>%
        arrange(desc(max_sust)) %>% 
        slice(1:20) %>%
        left_join(county_centers, by = "fips") %>%
        select(storm_id, name, max_sust) %>%
        mutate(storm_id = paste0(gsub("-", " (", storm_id), ")"),
               max_sust = round(max_sust)) %>%
        rename(Storm = storm_id,
               County = name,
               "Maximum sustained wind (m / s)" = max_sust) %>%
        xtable(caption = "Top storm-county pairs in terms of most intense sustained winds.", 
               label = "tab:top_wind")
print(a, include.rownames = FALSE, floating = TRUE,
      table.placement = "p", caption.placement = "top", 
      booktabs = TRUE, comment = FALSE, timestamp = NULL)
```

@@

```{r top_wind_map, echo = FALSE, fig.width = 8, fig.height = 6, fig.align = "center", fig.cap = "Number of storms for which each county experienced one of the top 2,000 intensities for distance"}
plot_top_x(wind_exposure, exposure_var = "max_sust",
           eastern_states = eastern_states, my_fips = my_fips,
           title = "Hits for top 2,000 Storm-County Pairs\nfor Wind Metric")
```

>>>>

```{r compare_rain_ex_counties, echo = FALSE, fig.width = 10, fig.height = 10, efig.align = "center", fig.cap = "Comparison of rain estimates for example counties."}
plot_county_rain_compare <- function(ex_fips, ex_dir, ex_title, 
                                     get_data = FALSE){
        if(get_data){
                write_daily_timeseries(ex_fips, coverage = 0,
                                       date_min = "1988-01-01",
                                       date_max = "2011-12-31",
                                       var = "PRCP",
                                       out_directory = ex_dir,
                                       keep_map = FALSE)
        }

@@

        ex_weather <- readRDS(paste0(ex_dir, "data/", ex_fips, ".rds"))

@@

```{r echo = FALSE, eval = FALSE}
allison_gauge <- read.csv("MWR_gauge_rain_data/allison_mwr.csv") 
allison_fips <- sprintf("%05s", as.character(unique(allison_gauge$fips)))
# allison_fips <- sample(allison_fips, 10)
date_range <- closest_dist %>% 
        filter(storm_id == "Allison-2001") %>%
        summarize(earliest_date = min(ymd(closest_date)) - ddays(2),
                  latest_date = max(ymd(closest_date)) + ddays(2))
write_daily_timeseries(allison_fips, coverage = 0,
                       date_min = format(date_range[1, 1], "%Y-%m-%d"),
                       date_max = format(date_range[1, 2], "%Y-%m-%d"),
                       var = "PRCP", out_directory = "county_ave/allison/",
                       keep_map = FALSE)

load("MWR_gauge_rain_data/frances_gauge.RData")
frances_fips <- sprintf("%05s", as.character(unique(frances_gauge$fips)))
# frances_fips <- sample(frances_fips, 10)
date_range <- closest_dist %>% 
        filter(storm_id == "Frances-2004") %>%
        summarize(earliest_date = min(ymd(closest_date)) - ddays(2),
                  latest_date = max(ymd(closest_date)) + ddays(2))
write_daily_timeseries(frances_fips, coverage = 0,
                       date_min = format(date_range[1, 1], "%Y-%m-%d"),
                       date_max = format(date_range[1, 2], "%Y-%m-%d"),
                       var = "PRCP", out_directory = "county_ave/frances/",
                       keep_map = FALSE)

load("MWR_gauge_rain_data/ike_gauge.RData")
ike_fips <- sprintf("%05s", as.character(unique(ike_gauge$fips)))
# ike_fips <- sample(ike_fips, 10)
date_range <- closest_dist %>% 
        filter(storm_id == "Ike-2008") %>%
        summarize(earliest_date = min(ymd(closest_date)) - ddays(2),
                  latest_date = max(ymd(closest_date)) + ddays(2))
write_daily_timeseries(ike_fips, coverage = 0,
                       date_min = format(date_range[1, 1], "%Y-%m-%d"),
                       date_max = format(date_range[1, 2], "%Y-%m-%d"),
                       var = "PRCP", out_directory = "county_ave/ike/",
                       keep_map = FALSE)

load("MWR_gauge_rain_data/alex_gauge.RData")
alex_fips <- sprintf("%05s", as.character(unique(alex_gauge$fips)))
# alex_fips <- sample(alex_fips, 10)
date_range <- closest_dist %>% 
        filter(storm_id == "Alex-2010") %>%
        summarize(earliest_date = min(ymd(closest_date)) - ddays(2),
                  latest_date = max(ymd(closest_date)) + ddays(2))
write_daily_timeseries(alex_fips, coverage = 0,
                       date_min = format(date_range[1, 1], "%Y-%m-%d"),
                       date_max = format(date_range[1, 2], "%Y-%m-%d"),
                       var = "PRCP", out_directory = "county_ave/alex/",
                       keep_map = FALSE)
```

```{r county_ave_comparison, echo = FALSE, fig.width = 10, fig.height = 10, fig.align = "center", fig.cap = "Comparison of county-level average cumulative rainfall for the period from two days before to one day after the storm's closest approach to a county, using NLDAS-2 versus ground-based monitors, for counties with both estimates available for four example storms. The size of each point indicates the average number of monitors reporting from that county during the window (larger size indicates more monitors). The text over each point gives the FIPS county identifier."}
plot_county_ave_compare <- function(county_dir, storm_id){
        storm_title <- paste0(paste(unlist(strsplit(
                as.character(storm_id), "-")),
                             collapse = " ("), ")")
        storm_year <- gsub(".+-", "", storm_id)
        counties <- gsub(".rds", "",
                         list.files(paste0("county_ave/", county_dir, "/data")))
        if(county_dir == "frances"){
                counties <- counties[counties != "51019"] # Missing all data
        }

@@

        for(x in counties){
                county_weather <- readRDS(paste0("county_ave/", county_dir, "/data/",
                                                 x, ".rds")) %>%
                        dplyr::mutate(date = ymd(date),
                                      prcp = prcp / 10) %>%
                        dplyr::mutate(fips = x)
                storm_ave[[x]] <- county_weather
                }
        storm_ave <- do.call("rbind", storm_ave)

@@

        compare_wind <- county_wind(counties = unique(data$fips),
                             start_year = storm_year, 
                             end_year = storm_year,
                             wind_limit = 0,
                             wind_var = "vmax_sust") %>%
                dplyr::filter(storm_id == data$storm_id[1]) %>%
                dplyr::right_join(data, by = "fips")

        county_range <- compare_wind %>% 
                group_by(fips) %>%
                summarize(lowest_wind = min(wind_sust),
                  highest_wind = max(wind_sust),
                  model_wind = mean(vmax_sust))

        plot_mwr <- ggplot(compare_wind, 
                   aes(x = vmax_sust, y = wind_sust)) + 
        geom_hline(yintercept = 15, color = "lightgray") + 
        geom_vline(xintercept = 15, color = "lightgray") + 
                geom_abline(aes(intercept = 0, slope = 1), color = "gray") + 
        geom_segment(data = county_range, 
                     aes(x = model_wind, xend = model_wind,
                         y = lowest_wind, yend = highest_wind,
                         group = fips)) + 
        geom_point(aes(group = NA), alpha = 0.5) + 
        ggthemes::theme_few() + 
        xlim(range(compare_wind$max_sust, compare_wind$wind_sust)) +
        ylim(range(compare_wind$max_sust, compare_wind$wind_sust)) + 
        ggtitle(storm_title) + 
        xlab("Modeled sustained wind (m / s)") + 
        ylab("Monthly Weather Report\n sustained wind reports (m / s)") + 
        scale_color_manual(values = c("TRUE" = "blue", "FALSE" = "black"), 
                           guide = "none") 
        return(plot_mwr)
}

@@

```{r top_wind, echo = FALSE, results = "asis"}
a <- wind_exposure %>%
        arrange(desc(vmax_sust)) %>% 
        slice(1:20) %>%
        left_join(county_centers, by = "fips") %>%
        select(storm_id, name, vmax_sust) %>%
        mutate(storm_id = paste0(gsub("-", " (", storm_id), ")"),
               max_sust = round(vmax_sust)) %>%
        rename(Storm = storm_id,
               County = name,
               "Maximum sustained wind (m / s)" = vmax_sust) %>%
        xtable(caption = "Top storm-county pairs in terms of most intense sustained winds.", 
               label = "tab:top_wind")
print(a, include.rownames = FALSE, floating = TRUE,
      table.placement = "p", caption.placement = "top", 
      booktabs = TRUE, comment = FALSE, timestamp = NULL)
```

@@

```{r top_wind_map, echo = FALSE, fig.width = 8, fig.height = 6, fig.align = "center", fig.cap = "Number of storms for which each county experienced one of the top 2,000 intensities for distance"}
plot_top_x(wind_exposure, exposure_var = "vmax_sust",
           eastern_states = eastern_states, my_fips = my_fips,
           title = "Hits for top 2,000 Storm-County Pairs\nfor Wind Metric")
```
